Generalized Bradley-Terry models

### 2025-03-10 inst/jss_hyper3_real.Rnw
more work

  The \code{hyper2} package provides functionality to work with
  extensions of the Bradley-Terry probability model such as reified
  entities (monsters) that account for known enhancements in
  competitive ability.  Here, I present a generalization of
  \code{hyper2} in which Bradley-Terry strengths may be enhanced using
  a multiplicative factor $\lambda$.  Generally, ordinary
  Bradley-Terry is recovered if $\lambda=1$.  This approach offers
  multiple advantages over reified Bradley-Terry, and several examples
  are given.
}
\Keywords{Plackett-Luce,
  Bradley-Terry, Mann-Whitney} \Plainkeywords{Plackett-Luce,
  Bradley-Terry, Mann-Whitney}

@@

Further, one might follow~\cite{hankin2017_rnw} and consider teams of
players.  In doubles tennis, for example, $\left\lbrace
p_1,p_2\right\rbrace$ might play $\left\lbrace p_3,p_4\right\rbrace$
and win with probability
$\frac{p_1+p_2}{p_1+p_2+p_3+p_4}$~\citep{hankin2010}.  It is natural to
combine this idea with that of a multiplicative factor.  For example,
suppose $p_1$ has a special move against $p_3$, and $p_4$ has just
been given a performance-enhancing drug, we might have a likelihood
functions proportional to $\frac{\lambda p_1+ p_2}{\lambda
  p_1+p_2+p_3+\mu p_4}$.  Then the effectiveness of $p_3$'s special
move and the enhancement afforded by the drug are measured by
$\lambda$ and $\mu$ respectively with values $>1$ corresponding to a
positive effect.

In this paper I present a generalization of the {\tt hyper2} suite for
Bradley-Terry likelihoods to accommodate general multiplicative
factors.  The most general likelihood function would be

\begin{equation}
  \prod_{i=1}^N
  \sum_{j=1}^n
 \left(\lambda_{ij}p_i\right)^{w_i}
  \end{equation}

where the $p_i$ are traditional Bradley-Terry strengths and the
$\lambda_{ij}\geqslant 0$ are multiplicative factors that account for
ephemeral or temporary changes in competitive ability.  In practice,
almost all the $\lambda_{ij}$ are zero.  To fix ideas, consider the
data presented in Table~\ref{oldfirm}.  Ignoring draws for the moment,
we have a likelihood proportional to

@@

\begin{equation}
  \left(\frac{\lambda p_R}{\lambda p_R + p_C}\right)^{81}\cdot
  \left(\frac{\lambda p_R}{\lambda p_R + p_F}\right)^{69}\cdots
  \left(\frac{\lambda p_L}{\lambda p_L + p_F}\right)^{81}\cdots
  \left(\frac{p_C}{p_C + \lambda p_R}\right)^{42}\cdots
    \left(\frac{p_F}{p_F + \lambda p_L}\right)^{9}
\end{equation}

In {\tt hyper3} idiom, a term such as $\lambda p_R + p_C$ is
represented as a named numeric vector.  For example, if the home
strength were $\lambda = 1.8$ then the first bracket would be {\tt
  c(Rangers=1.8, Celtic=1)} with an associated power of $-81$.  The
computational methodology is a generalization of that used
in~\cite{hankin2024_hyper3} in which the multiplicative factors are
known non-negative integers.

\subsection{The package in use}
Package idiom for creating a likelihood function for this dataset
would look something like this: Considering only Rangers and Celtic:
Rangers at home:

<<defineoldfirm, cache=TRUE>>=
oldfirm <- function(lambda){
    H <- hyper3(pnames = c("Rangers", "Celtic"))
    H[c(Rangers = lambda)            ] %<>% add(81)
    H[c(Celtic = 1)                  ] %<>% add(42)
    H[c(Rangers = lambda, Celtic = 1)] %<>% subtract(81 + 42)

    H[c(Celtic = lambda)             ] %<>% add(76)
    H[c(Rangers = 1)                 ] %<>% add(46)
    H[c(Celtic = lambda, Rangers = 1)] %<>% subtract(46 + 76)
    return(H)
}
@

@@

We see a log-likelihood function for all $82+42+76+46=245$
Rangers-Celtic matches that were not score draws, conditional on
assumed value $\lambda=1.88$.  We may make inferences about the true
value of $\lambda$ using profile support:

@@

Thus {\tt f()} returns the support at the evaluate, conditional on a
supplied value of $\lambda$.  We may use this to create a profile
likelihood:

@@

@@

We are now in a position to consider the entirety of
Table~\ref{oldfirm}.  Bespoke function \code{home\_draw\_away3()}
converts data in the form of Table~\ref{oldfirm} to a \code{hyper3}
likelihood function following Table~\ref{windrawloss}; it takes
numerical values for the home advantage $\lambda$ and proclivity to
draw $D$:

@@

Such estimates have rich predictive power.  For example, if Rangers
were to play Celtic with both teams away [thus nullifying the home
  ground advantage: $\lambda=1$ in Table~\ref{windrawloss}], the
probability of win/draw/loss for Rangers would be
$\left(\frac{p_R}{(1+D)(p_R+p_C)};\frac{D}{1+D};\frac{p_R}{(1+D)(p_R+p_C)}\right)$,
estimated at $(0.39;0.27;0.34)$.

<<makecontourdata, cache=TRUE>>=
n <- 20
lambda <- seq(from=1, to=3, len=n)
D <- seq(from=0.1, to=0.6, len=n)
V <- as.matrix(expand.grid(lambda, D))
LL <- apply(V, 1, f)
LL <- pmax(-20, LL - max(LL))
@

<<>>=
lambda
@

@@

In the following, I use the ``race'' metaphor: five runners take part
in a race and arrive in order: $a\succ b\succ c\succ d\succ e$.
Consider the following Plackett-Luce likelihood function for this
observation:

@@

Note the use of multiplication between the terms, indicating
conditional independence.  We may use the multiplicative
generalization of Bradley-Terry strengths to introduce an element of
non-independence between the terms.  There are many ways to do this,
but one simple case would be to define equivalence classes of the
competitors with each equivalence class comprising mutually supporting
runners.  Equivalence classes of size one correspond to unsupported
runners.  The metaphor would be that a runner who has finished the
race is able to support other members of his equivalence class by
cheering his teammates, boosting their performance.  The \pkg{hyper2}
package includes function \code{cheering()} which implements this
functionality.

@@

The classic red bus-blue bus problem was originally stated by
\cite{mcfadden1980} although different metaphors have been employed
for the same phenomenon by \cite{tverskey1972}.  The canonical problem
is as follows.  A traveller has a choice of going to work by different
transport methods.  Here, we can assume that the traveller has strong
but unknown views on different forms of transport which might
incorporate a random component.  If considering whether to use a car
or a bus, the colour of the bus is of no importance to him.  Such
preferences are problematic for Bradley-Terry: If our subject is
indifferent between cars and buses, say, $p_{C} = p_{RB}$ and also
indifferent to colour of bus, so $p_{RB}=p_{BB}$, the following
paradox occurs.  Bradley-Terry strengths must be equal so $p_{C} =
p_{RB} = p_{BB}=\frac{1}{3}$.  Thus $\Prob(C| \left\lbrace
C,RB\right\rbrace)=\frac{1}{2}$.  Now we introduce a blue bus as an
option.  Any realistic choice model will have $\Prob(C | \left\lbrace
C,RB,BB\right\rbrace)=\frac{1}{2}$ but Bradley-Terry gives us
$\Prob(C| \left\lbrace C,RB\right\rbrace)=\frac{1}{3}$.  Various
solutions to this problem have been proposed.  \cite{tverskey1972},
for example, presents an approach involving overlapping ``aspects'' of
the elements of choice set; \cite{mcfadden1980} suggests using a
generalized extreme-value model.

@@

Above, we see from the function body that he reported $RB\succ BB\succ
C\succ T\succ W$ three times [first row], and so on; perhaps his
ranking depends on the weather or how tired he is on any given day.
Observe that in almost every case he ranks RB and BB consecutively.
Function \code{RB_BB_LF()} takes argument \code{lambda} that
quantifies the perceived similarity between RB and BB in the same way
as function \code{oldfirm()} above.

@@

We can thus improve the hypothesis that $\lambda=1$, corresponding to
an unaltered Plackett-Luce system, by over 5.6 units of support by
dint of choosing instead $\lambda = 10.49$.  This easily satisfies
\citeauthor{edwards1992}'s two-units-of-support criterion;
alternatively, we could observe that this observation lies in the tail
region of its asymptotic $\chi^2_1$ distribution (Wilks's theorem),
with a $p$-value of \Sexpr{signif(pchisq(suppdiff*2, df=1,
  lower.tail=FALSE),3)}.  It is straightforward to find a profile
support function for $\lambda$, Figure~\ref{profilesupportlambdaRBBB}.
We see that support of $\log\lambda$ is closely quadratic over a broad
range.

\begin{figure}[htbp]
\begin{center}
<<plotproflike>>=
lambda <- exp(seq(from=log(0.9), to=log(67), len=19)) 
L <- sapply(lambda, o)
plot(log(lambda), L-max(L), type="b")
abline(h=c(0, -2))
abline(v=0)
@ 
\caption{A\label{profilesupportlambdaRBBB} profile support function
  for the red bus, blue bus parameter $\lambda$, synthetic dataset.
  Vertical line indicates null of $\lambda=1$, horizontal lines show
  maximum support and $-2$}
\end{center}
\end{figure}

@@

<<showsushitable>>=
head(noquote(sushi_table))
@

@@

so class 2 is the tuna group and class 1 everything else.  We can now
use these datapoint to define a bespoke likelihood function, using
$\lambda$ as the argument:

@@

\begin{figure}[htbp]
  \begin{center}
<<plotlikesushi>>=
plot(log(s),L-max(L),type='b',
     xlab=expression(log(lambda)), ylab='Log-likelihood')
@ 
\caption{A\label{profilesupportsushi} profile support function for the
  exchangeability term $\lambda$ in the case of sushi preferences.
  Vertical gray line shows $H_0\colon\lambda =1$, horizontal lines
  show the two-units-of-support limits}
\end{center}
\end{figure}

@@

From the first line we see that {\em Cloud9} played {\em FaZe Clan} on
the {\em inferno} map, won the match, and \skad\ had the least number
of deaths, followed by \rush, then \tarik, and so on.  Consider two
{\em Cloud9} players: \skad\ and \stew.  These teammates, although
comparable in skill [we fail to reject $H_0\colon p_{\mbox{\tiny\sc
      Skadoodle}}\neq p_{\mbox{\tiny\sc Stewie2k}}$, Plackett-Luce,
  $p=0.28$] have playing styles that stand in dramatic contrast.
Commentators characterise \skad\ as emphasising reactive map control
while \stew\ is more of a disruptive presence engaging in
high-variance play.

@@

<<findopt, cache=TRUE, echo=TRUE>>=
optlam <- optimize(f,c(0.09, 0.14), maximum = TRUE)
evaluate <- optlam$maximum
support <- optlam$objective - f(1)
pvalue <- pchisq(support, df = 1, lower.tail = FALSE)
@

<<calcklikelambda, cache=TRUE, echo=FALSE>>=
lam <- exp(seq(from = -4,to = log(1.1), len = 23))
L <- sapply(lam, f)
@

<<findinterval, cache=TRUE, echo=FALSE>>=
g <- function(l){f(l) - optlam$objective  + 2}
jj1 <- uniroot(g, interval = exp(c(-4,-3)))
jj2 <- uniroot(g, interval = exp(c(-1,-0)))
lowerlimit <- jj1$root
upperlimit <- jj2$root
@

\begin{figure}[htbp]
  \begin{center}
<<plotlikelam, echo=FALSE>>=
plot(log(lam), L - max(L), type='b')
segments(x0=log(optlam$maximum),y0 = -0.5, y1 = 0.1,col='red')
abline(h=c(0,-2), col = 'gray')
segments(x0 = log(lowerlimit), y0 = -2.2, y1 = -1.8, col = 'blue')
segments(x0 = log(upperlimit), y0 = -2.2, y1 = -1.8, col = 'blue')
abline(v = 0, lty = 2)
@ 
\caption{A\label{profsupcounterstrike} profile support function for
  the red bus-blue bus parameter $\log\lambda$ applied to mutually
  supporting equivalence class
  $\left\lbrace\mbox{\skad},\mbox{\stew}\right\rbrace$.  Vertical gray
  line shows $H_0\colon\lambda =1$, for which the support is about
  $\Sexpr{round(support,2)}$ relative to the evaluate
  $\hat{\lambda}\simeq\Sexpr{round(evaluate,3)}$ (small red bar),
  corresponding to a $p$-value of about $\Sexpr{round(pvalue,3)}$.
  Horizontal lines show the two-units-of-support credible interval of
  $(\Sexpr{round(lowerlimit,3)},\Sexpr{round(upperlimit,3)})$, marked
  in blue}
\end{center}
\end{figure}

To assess this, we follow exactly the same protocol as for the sushi
preference dataset but using the deathorder statistic from the dataset
above.  We specify two equivalence classes: one comprising \skad\ and
\stew, and one everyone else.  Figure~\ref{profsupcounterstrike} gives
a support function for $\lambda$ showing, among other things, that we
may reject the hypothesis that $\lambda=1$ in favour of $\lambda<1$ by
a likelihood ratio of about $\Sexpr{round(support,2)}$.  We observe
that this exceeds the two units of support criterion of
\cite{edwards1992} [asymptotic $p$-value
  $\simeq\Sexpr{round(pvalue,3)}$] so we may infer that there is
indeed some systematic mechanism---perhaps a positive feedback
loop---which tends to magnify differences in death statistic.

@@

Plackett-Luce likelihood functions for rank datasets have been

>>>>

  A generalization of \code{hyper2} is presented in which
  Bradley-Terry strengths may be modified using a multiplicative
  factor $\lambda$; Generally, ordinary Bradley-Terry is recovered if
  $\lambda=1$.  This approach offers multiple advantages over other
  generalizations such as reified Bradley-Terry.  It becomes possible
  to represent non-independence of successive observations in
  Plackett-Luce likelihood functions which allows a parametrization of
  the ``red bus, blue bus'' problem in preference analysis.  I present
  examples drawn from the fields of food preference and the e-sport
  {\em Counterstrike}.  }

\Keywords{Plackett-Luce, Bradley-Terry, Mann-Whitney}
\Plainkeywords{Plackett-Luce, Bradley-Terry, Mann-Whitney}

@@

Further, one might follow~\cite{hankin2017_rnw} and consider teams of
players.  In doubles tennis, for example, $\left\lbrace
p_1,p_2\right\rbrace$ might play $\left\lbrace p_3,p_4\right\rbrace$
and win with probability
$\frac{p_1+p_2}{p_1+p_2+p_3+p_4}$~\citep{hankin2010}.  It is natural
to combine this idea with that of a multiplicative factor.  For
example, suppose $p_1$ has a special move against $p_3$, and $p_4$ has
just been given a performance-enhancing drug.  If so, we might have
likelihood function

\begin{equation}\label{like1234}
  \mathcal{L}{\left(p_1,p_2,p_3,p_4;\lambda,\mu\right)}\propto
  \frac{
    \left(\lambda p_1 + p_2\right)^{w_{12}}\cdot
    \left(p_3 + \mu p_4\right)^{w_{34}}
    }{
    \left(\lambda p_1+p_2+p_3+\mu p_4\right)^{w_{12} + w_{34}}
    }
\end{equation}

where $w_{12}$ and $w_{34}$ are the number of wins by each team
respectively.  Then $\lambda$ and $\mu$ represent the effectiveness of
$p_1$'s special move and the enhancement afforded to $p_4$ by the
drug, with values $>1$ corresponding to a positive effect.  In this
paper I generalize the {\tt hyper2} suite for Bradley-Terry
likelihoods~\citep{hankin2017_rnw} to accommodate this class of
multiplicative factors.  The most general likelihood function would be
proportional to

\begin{equation}
  \prod_{i=1}^N
  \sum_{j=1}^n
 \left(\lambda_{ij}p_j\right)^{w_i}
  \end{equation}

where the $p_j$ are traditional Bradley-Terry strengths, $p_j\geqslant
0$, $\sum p_j=1$, and the $\lambda_{ij}\geqslant 0$ are multiplicative
factors that account for ephemeral or temporary changes in competitive
ability.  The data $w_1,\ldots w_N$ correspond to the number of
``wins'' and number of matches, as per $w_{12},w_{34}$ in
equation~\ref{like1234}.  In practice, almost all the $\lambda_{ij}$
are zero.  To fix ideas, consider the data presented in
Table~\ref{oldfirm}.  Ignoring draws for the moment, we have a
likelihood proportional to

@@

\begin{equation}\label{firstfewRCFL}
  \left(\frac{\lambda p_R}{\lambda p_R + p_C}\right)^{81}\cdot
  \left(\frac{\lambda p_R}{\lambda p_R + p_F}\right)^{69}\cdots
  \left(\frac{\lambda p_L}{\lambda p_L + p_F}\right)^{81}\cdots
  \left(\frac{p_C}{p_C + \lambda p_R}\right)^{42}\cdots
    \left(\frac{p_F}{p_F + \lambda p_L}\right)^{9}
\end{equation}

where the home-ground advantage is represented by~$\lambda$.  In {\tt
  hyper3} idiom, a term such as $\lambda p_R + p_C$ is represented as
a named numeric vector.  For example, if the home strength were
$\lambda = 1.8$ then the first bracket would be {\tt c(Rangers=1.8,
  Celtic=1)} with an associated power of $-81$.  The computational
methodology is a generalization of that used
in~\cite{hankin2024_hyper3} in which the multiplicative factors
correspond to number of clones.  However, in~\cite{hankin2024_hyper3}
the factors are known non-negative integers, and here they are, in
general, unknown non-negative real numbers.

\subsection{The package in use}
Package idiom for creating a likelihood function for this dataset is
straightforward.  Considering first only Rangers-Celtic matches:

<<defineoldfirm, cache=TRUE>>=
oldfirm <- function(lambda){
    H <- hyper3(pnames = c("Rangers", "Celtic"))
    H[c(Rangers = lambda            )] %<>%      add(81     )
    H[c(                  Celtic = 1)] %<>%      add(     42)
    H[c(Rangers = lambda, Celtic = 1)] %<>% subtract(81 + 42)

    H[c(Celtic = lambda             )] %<>%      add(     76)
    H[c(                 Rangers = 1)] %<>%      add(46     )
    H[c(Celtic = lambda, Rangers = 1)] %<>% subtract(46 + 76)
    return(H)
}
@

@@

We see a log-likelihood function for all $82+42+76+46=245$
Rangers-Celtic matches that were not score draws, conditional on
assumed value $\lambda=1.88$.  This corresponds to the relevant terms
of equation~\ref{firstfewRCFL}.  We may make inferences about the true
value of $\lambda$ using profile support:

@@

Thus {\tt f()} returns the support at the evaluate, conditional on a
supplied value of $\lambda$.  We may use this to create a profile
likelihood (figure~\ref{profilesupportlambda}), or to find the maximum
likelihood estimate for $\lambda$:

@@

Then using this value we may estimate the Bradley-Terry strengths of
Rangers and Celtic:

@@

We are now in a position to consider the entirety of
Table~\ref{oldfirm}.  Bespoke package function
\code{home\_draw\_away3()} converts data in the form of
Table~\ref{oldfirm} to a \code{hyper3} likelihood function following
Table~\ref{windrawloss}; it takes numerical values for the home
advantage $\lambda$ and proclivity to draw $D$:

@@

Such estimates have rich predictive power.  For example, if Rangers
were to play Celtic with both teams away [thus nullifying the home
  ground advantage: $\lambda=1$ in Table~\ref{windrawloss}], the
probability of win/draw/loss for Rangers would be
$\left(\frac{p_R}{(1+D)(p_R+p_C)};\frac{D}{1+D};\frac{p_C}{(1+D)(p_R+p_C)}\right)$,
estimated at $(0.39;0.27;0.34)$.

<<makecontourdata, echo=FALSE, cache=TRUE>>=
n <- 20
lambda <- seq(from=1, to=3, len=n)
D <- seq(from=0.1, to=0.6, len=n)
V <- as.matrix(expand.grid(lambda, D))
LL <- apply(V, 1, f)
LL <- pmax(-20, LL - max(LL))
@

@@

In the following, I use the ``race'' metaphor: five runners take part
in a race and arrive in order $a\succ b\succ c\succ d\succ e$.  Thus
competitor $a$ wins, $b$ comes second, and so on.  Consider the
following Plackett-Luce likelihood function for this observation:

@@

Note the use of multiplication between the terms, indicating
conditional independence.  Now we use the multiplicative
generalization of Bradley-Terry strengths to introduce an element of
non-independence between the terms.  There are many ways to do this,
but one simple way would be to define equivalence classes of the
competitors with each equivalence class comprising mutually supporting
runners.  The metaphor would be that a runner who has finished the
race is able to support other members of his equivalence class by
cheering his still-running teammates, boosting their performance.  The
\pkg{hyper2} package includes function \code{cheering()} which
implements this functionality.

@@

The classic red bus-blue bus problem was originally stated by
\cite{mcfadden1980} although different metaphors have been employed
for the same phenomenon by \cite{tverskey1972}.  The canonical problem
is as follows.  A commuter has a choice of going to work by different
transport methods.  Here, we can assume that he has possibly
strong---but unknown---views on different forms of transport which
might incorporate a random component.  If considering whether to use a
car or a bus, the colour of the bus is of no importance to him.  Such
preferences are problematic for Bradley-Terry: If our subject is
indifferent between cars and buses, then $p_{C} = p_{RB}$.  If he is
also indifferent to colour of bus, so $p_{RB}=p_{BB}$, the following
paradox occurs.  Bradley-Terry strengths must be equal so $p_{C} =
p_{RB} = p_{BB}=\frac{1}{3}$.  Thus $\Prob(C| \left\lbrace
C,RB\right\rbrace)=\frac{1}{2}$.  Now we introduce a blue bus as an
option.  Any realistic choice model will have $\Prob(C | \left\lbrace
C,RB,BB\right\rbrace)=\frac{1}{2}$ but Bradley-Terry gives us
$\frac{1}{3}$.  Various solutions to this problem have been proposed.
\cite{tverskey1972}, for example, presents an approach involving
overlapping ``aspects'' of the elements of choice set;
\cite{mcfadden1980} suggests using a generalized extreme-value model.

@@

Above, we see from the function body that he reported $RB\succ BB\succ
C\succ T\succ W$ three times [first row], and so on; perhaps his
ranking depends on the weather or how tired he is on any given day.
Vector \code{ec} places the options into equivalence classes;
equivalence class 3 comprises the two buses and the other classes are
unsupported as they are of size one.  Vector \code{h} encodes the
strength of the nonindependence.  Observe that in almost every case he
ranks RB and BB consecutively.  Function \code{RB_BB_LF()} takes
argument \code{lambda} that quantifies the perceived similarity
between RB and BB in the same way as function \code{oldfirm()} above.

@@

We can thus improve the hypothesis that $\lambda=1$, corresponding to
an unaltered Plackett-Luce system, by over 5.6 units of support by
dint of choosing instead $\lambda = 10.49$.  This easily satisfies
\citeauthor{edwards1992}'s two-units-of-support criterion;
alternatively, we could observe that this observation lies in the tail
region of its asymptotic $\chi^2_1$ distribution (Wilks's theorem),
with a $p$-value of \Sexpr{signif(pchisq(suppdiff*2, df=1,
  lower.tail=FALSE),3)}.

%It is straightforward to find a profile
%support function for $\lambda$, Figure~\ref{profilesupportlambdaRBBB}.
%We see that support of $\log\lambda$ is closely quadratic over a broad
%range.

%\begin{figure}[htbp]
%\begin{center}
<<plotproflike, echo=FALSE>>=
if(FALSE){
    lambda <- exp(seq(from=log(0.9), to=log(67), len=19)) 
    L <- sapply(lambda, o)
    plot(log(lambda), L-max(L), type="b")
    abline(h=c(0, -2))
    abline(v=0)
}
@ 
%\caption{A\label{profilesupportlambdaRBBB} profile support function
%  for the red bus, blue bus parameter $\lambda$, synthetic dataset.
%  Vertical line indicates null of $\lambda=1$, horizontal lines show
%  maximum support and $-2$}
%\end{center}
%\end{figure}

@@

<<showsushitable>>=
noquote(head(sushi_table))
@

@@

so class 2 is the tuna group and class 1 everything else.  We can now
define a bespoke likelihood function for these observations, using
$\lambda$ as the argument:

@@

<<findstuffsushi,echo=FALSE,cache=TRUE>>=
osushi <- optimize(f, c(2,4), maximum=TRUE)
@

<<tinysushi,echo=FALSE>>=
sushi_maxlambda <- osushi$maximum
@

<<findothersushistuff, echo=FALSE, cache=TRUE>>=
g <- function(lambda){f(lambda) - osushi$objective + 2}
sushi_lower <- uniroot(g, interval = exp(c(0,1)), tol = 0.01)$root
sushi_upper <- uniroot(g, interval = exp(c(1,2)), tol = 0.01)$root
@

\begin{figure}[htbp]
  \begin{center}
<<plotlikesushi, echo=FALSE>>=
plot(log(s),L-max(L),type='b',
     xlab=expression(log(lambda)), ylab='Log-likelihood')
abline(v=0,col='gray')
abline(h=c(0,-2),col='gray')
segments(x0 = log(sushi_maxlambda), y0 =  0.1, y1 = -1.0, col = 'red' )
segments(x0 = log(sushi_lower    ), y0 = -2.2, y1 = -1.8, col = 'blue')
segments(x0 = log(sushi_upper    ), y0 = -2.2, y1 = -1.8, col = 'blue')
@ 
\caption{A\label{profilesupportsushi} profile support function for the
  exchangeability term $\lambda$ in the case of sushi preferences.
  Vertical gray line shows $H_0\colon\lambda =1$, horizontal lines
  show the two-units-of-support limits with the credible interval of
  $(\Sexpr{round(sushi_lower,2)},\Sexpr{round(sushi_upper,2)})$,
  marked in blue.  Red bar shows evaluate at
  $\lambda\simeq\Sexpr{round(sushi_maxlambda,2)}$}
\end{center}
\end{figure}

@@

From the first line we see that {\em Cloud9} played {\em FaZe Clan} on
the {\em inferno} map, won the match, and \skad\ had the least number
of deaths, followed by \rush, then \tarik, and so on.  Consider two
{\em Cloud9} players: \skad\ and \stew.  These teammates, although
comparable in skill [we fail to reject $H_0\colon p_{\mbox{\tiny\sc
      Skadoodle}} = p_{\mbox{\tiny\sc Stewie2k}}$, Plackett-Luce,
  $p=0.28$] have playing styles that stand in dramatic contrast.
Commentators characterise \skad\ as emphasising reactive map control
while \stew\ is more of a disruptive presence engaging in
high-variance play.

@@

<<csgo_findopt, cache=TRUE, echo=FALSE>>=
csgo_optimize <- optimize(f,c(0.09, 0.14), maximum = TRUE, tol = 0.01)
csgo_optlam   <- csgo_optimize$maximum
csgo_support  <- csgo_optimize$objective - f(1)
csgo_pvalue   <- pchisq(csgo_support, df = 1, lower.tail = FALSE)
@

<<csgo_calcklikelambda, cache=TRUE, echo=FALSE>>=
lam <- exp(seq(from = -4, to = log(1.1), len = 23))
L <- sapply(lam, f)
@

<<cgso_findinterval, cache=TRUE, echo=FALSE>>=
g <- function(l){f(l) - csgo_optimize$objective  + 2}
csgo_lower <- uniroot(g, interval = exp(c(-4,-3)))$root
csgo_upper <- uniroot(g, interval = exp(c(-1,-0)))$root
@

\begin{figure}[htbp]
  \begin{center}
<<csgo_plotlikelam, echo=FALSE>>=
plot(log(lam), L - max(L), type='b')
segments(x0 = log(csgo_optlam),y0 = -0.5, y1 =  0.1, col='red')
segments(x0 = log(csgo_lower), y0 = -2.2, y1 = -1.8, col = 'blue')
segments(x0 = log(csgo_upper), y0 = -2.2, y1 = -1.8, col = 'blue')
abline(v = 0, col = 'gray')
abline(h = c(0, -2), col = 'gray')
@ 
\caption{A\label{profsupcounterstrike} profile support function for
  the red bus-blue bus parameter $\log\lambda$ applied to mutually
  supporting equivalence class
  $\left\lbrace\mbox{\skad},\mbox{\stew}\right\rbrace$.  Vertical gray
  line shows $H_0\colon\lambda =1$, for which the support is about
  $\Sexpr{round(csgo_support,2)}$ relative to the evaluate
  $\hat{\lambda}\simeq\Sexpr{round(csgo_optlam,3)}$ (small red bar),
  corresponding to a $p$-value of about
  $\Sexpr{round(csgo_pvalue,3)}$.  Horizontal lines show the
  two-units-of-support credible interval of
  $(\Sexpr{round(csgo_lower,3)},\Sexpr{round(csgo_upper,3)})$, marked
  in blue}
\end{center}
\end{figure}

To assess this, we follow exactly the same protocol as for the sushi
preference dataset but using the deathorder statistic from the dataset
above.  We specify two equivalence classes: one comprising \skad\ and
\stew, and one everyone else.  Figure~\ref{profsupcounterstrike} gives
a support function for $\lambda$ showing, among other things, that we
may reject the hypothesis that $\lambda=1$ in favour of $\lambda<1$ by
a likelihood ratio of about $\Sexpr{round(csgo_support,2)}$.  We
observe that this exceeds the two units of support criterion of
\cite{edwards1992} [asymptotic $p$-value
  $\simeq\Sexpr{round(csgo_pvalue,3)}$] so we may infer that there is
indeed some systematic mechanism---perhaps a positive feedback
loop---which tends to magnify differences in death statistic.

@@

Bradley-Terry strengths $p_1,\ldots,p_n$, $p_i\geqslant 0$, $\sum
p_i=1$ are generalized with non-negative multiplicative $\lambda$-
factors that represent temporary or ephemeral changes in competitive
strength.  The {\tt hyper2} R package implements the resulting
likelihood functions, and furnishes a wide range of functionality
geared towards assessing meaningful nulls using likelihood methods.

The standard Plackett-Luce likelihood function
$\prod_{i=1}^{n}p_{(i)}\left[\sum_{j=i}^n p_{(j)}\right]^{-1}$ over
Bradley-Terry strengths assumes conditional independence of successive
rank observations $p_{(1)},\ldots p_{(n)}$, an assumption which is
broken in the current work.  As indicated above, multiplicative
$\lambda$- factors may be used to implement many different
non-independence schemes and comparing these might be interesting.

Further work might include an individualised multiplicative draw
factor, giving rise to likelihood functions such as
$\frac{p_1^{w_1};(\lambda_1p_1+\lambda_2p_2)^{d_{12}};
  p_2^{w_2}}{\left(p_1(1+\lambda_1)+p_2(1+\lambda_2)\right)^{w_1+d_{12}+w_2}}$.
One might also consider multiplicative factors for the strengths of
{\em reified} entities.  These modifications would be directly
representable with existing {\tt hyper3} idiom.
